---
title: "ps2"
author: "Cali Li"
date: "10/11/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data setup
```{r, echo=FALSE}
# load packages:
library(tidyverse)
library(lme4)
library(lmerTest)
library(mousetrap)
library(kableExtra) 

# read or load data: -----------------------------------------------------------

recs = readr::read_delim('C:/Users/wenji/Downloads/recs2015_public_v4.csv',
                             delim=',')

```

## Question 1.a

```{r, echo=FALSE}
# Question 1.a: ----------------------------------------------------------------
# the national average home temperature at night in winter
# (among homes that use space heating).
# variable in data: TEMPHOME == 1 ---> use space heating
#                   TEMPNITE != -2 ---> Winter temperature at night

# Mean by final & replicate weights: -------------------------------------------
mean_replicate = recs %>%
  filter(HEATHOME == 1 & TEMPNITE != -2) %>%
  gather(key = 'replicate', value = 'BRRWT', BRRWT1:BRRWT96) %>%
  select(replicate, BRRWT, TEMPNITE, NWEIGHT) %>%
  group_by(replicate) %>%
  mutate( means_rep = TEMPNITE * BRRWT / sum(BRRWT),
          means_tot = TEMPNITE * NWEIGHT / sum(NWEIGHT)) %>%
  ungroup() %>%
  summarize( std_err = sqrt( sum( {means_rep - means_tot}^2 ) *4/96 ),
             lwr = sum(means_tot) - qnorm(.975)*std_err,
             upr = sum(means_tot) + qnorm(.975)*std_err) %>%
  select(-std_err, means_tot)
mean_replicate
```


## Question 1.c

```{r, echo = FALSE}
decode_division = function(x){
  # Decodes numeric codes to division labels
  #
  # Args: 
  #   x: a single numeric code for a DIVISION
  #
  # Returns: The division label
  
  # Throw an error if x isn't numeric
  if(!is.numeric(x) | x < 1) stop('decode_division expects numeric input 
                                  indexed from 1!')
  
  switch(x,
         "NewEng", "MidAtlan", "ENCen",
         "WNCentral", "SoutAtlan", "ESCen",
         "WSCentral", "MountNth", "MountSth",
         "Pacific"
  )
}
decode_all_division = function(x){
  # Vectorizes decode_division above
  #
  # Args: 
  #  x: a vector of integer-valued division
  #
  # Returns: A vector of division labes or a "list" if some are unmatched.
  sapply(x, decode_division)
}

## urban type
decode_uatyp = function(x){
  
  if(x != "U" & x!= "C" & x != "R") stop('Invaild input for decode_uatyp!')
  
  switch(x,
         "U" = "UrbanArea",
         "C" = "UrbanCluster",
         "R" = "Rural"
  )
}
decode_all_uatyp = function(x){
  # Vectorizes decode_uatyp above
  #
  # Args: 
  #  x: a vector of integer-valued uatyp
  #
  # Returns: A vector of uatyp labes or a "list" if some are unmatched.
  sapply(x, decode_uatyp)
}

## fuelheat
decode_fuel = function(x){
  if(!is.numeric(x)) stop('decode_fuel expects numeric input!')
  x = as.character(x)
  switch(x,
         "1" = "NaturalGas",
         "2" = "Propane",
         "3" = "OilKerosene",
         "5" = "Electricity",
         "7" = "Wood",
         "21" = "OtherFuel",
         "-2" = "NotApplicable"
  )
}
decode_all_fuel = function(x){
  # Vectorizes decode_fuel above
  #
  # Args: 
  #  x: a vector of integer-valued heatfuel
  #
  # Returns: A vector of fuel labes or a "list" if some are unmatched.
  sapply(x, decode_fuel)
}

temp_nweights = recs %>% 
  filter(TEMPNITE >= 0 , TEMPGONE >= 0 , TEMPHOME >= 0) %>%
  mutate(division = decode_all_division(DIVISION), 
         uatyp = decode_all_uatyp(UATYP10)) %>%
  select(division, uatyp, tempnite = TEMPNITE, tempgone = TEMPGONE, 
         temphome = TEMPHOME, weight = NWEIGHT) %>%
  group_by(division, uatyp) %>%
  summarize( w4tempnite = sum(tempnite*weight) / sum(weight),
             w4tempgone = sum(tempgone*weight) / sum(weight),
             w4temphome = sum(temphome*weight) / sum(weight)
              )

# Mean by replicate weights & plot:
temp_replicate = recs %>%
  filter(TEMPNITE >= 0 , TEMPGONE >= 0 , TEMPHOME >= 0) %>%
  tidyr::pivot_longer(cols = starts_with('BRRWT'),
                      names_to = 'replicate',
                      values_to = 'BRRWT') %>%
  mutate(division = decode_all_division(DIVISION), 
         uatyp = decode_all_uatyp(UATYP10)) %>%
  select(division, uatyp, tempnite = TEMPNITE, tempgone = TEMPGONE, 
         temphome = TEMPHOME, weight = NWEIGHT,replicate, BRRWT) %>%
  replace_na(list(weight = 0)) %>%
  group_by(division, uatyp,replicate) %>%
  summarize(tempnite_rep = sum(tempnite * BRRWT) / sum(BRRWT),
            tempgone_rep = sum(tempgone * BRRWT) / sum(BRRWT),
            temphome_rep = sum(temphome * BRRWT) / sum(BRRWT)
             ) %>%
  ungroup() %>%
  left_join(temp_nweights, by = c('division', 'uatyp')) %>%
  group_by(division, uatyp) %>%
  summarize(w4tempnite = w4tempnite[1],                      # Compute se.
            w4tempgone = w4tempgone[1],
            w4temphome = w4temphome[1],
            std_err4tempnite = sqrt( sum( {tempnite_rep - w4tempnite}^2 )*4/96 ),
            std_err4tempgone = sqrt( sum( {tempgone_rep - w4tempgone}^2 )*4/96 ),
            std_err4temphome = sqrt( sum( {temphome_rep - w4temphome}^2 )*4/96 )
            ) %>%
  mutate(ci4tempnite = qnorm(.975)*std_err4tempnite,
         ci4tempgone = qnorm(.975)*std_err4tempgone,
         ci4temphome = qnorm(.975)*std_err4temphome
        ) %>%
  select(-std_err4tempnite, -std_err4temphome, -std_err4tempgone) %>%
  tidyr::pivot_longer( cols = starts_with('w4'),
                       names_to = 'temp_type',
                       values_to = 'mean_temp',
                       ) %>%
  mutate(ci = case_when(temp_type == 'w4tempnite' ~ ci4tempnite,
                        temp_type == 'w4tempgone' ~ ci4tempgone,
                        temp_type == 'w4temphome' ~ ci4temphome)) %>%
  ggplot(aes(x = uatyp, y = mean_temp, fill = temp_type)) +
  geom_bar(stat = "identity", position = 'dodge') +
  geom_errorbar(aes(ymin = mean_temp - ci, ymax = mean_temp + ci), 
                width = 0.5, position = position_dodge(0.9)) +
  facet_wrap(~ division, nrow = 5) + xlab('Urban Type') + 
  ylab('Temperature') + 
  ggtitle('Average Winter Home Temperatures') +
  theme_light()
temp_replicate
```

## Question 1.d

```{r, echo = FALSE}
decode_equipmuse = function(x) {
  if(!as.numeric(x)) stop('Invaild input for decode_equipmuse!')
  x = as.character(x)
  switch(x,
         "1" = "Set one temperature",
         "2" = "Manually adjust it at night or go out",
         "3" = "Automatically adjust the temperature ",
         "4" = "As needed",
         "5" = "Does not control it",
         "9" = "Other"
  )
}
decode_all_equipmuse = function(x) {
  # Vectorizes decode_equipmuse above
  #
  # Args: 
  #  x: a vector of integer-valued equipmuse
  #
  # Returns: A vector of equipmuse labes or a "list" if some are unmatched.
  sapply(x, decode_equipmuse)
}

## Function median:
my_median = function(x,y) {
  x = sort(x)
  i = 1
  cumu = 0
  while (i <= length(x)) {
    cumu = cumu + x[i]
    if (cumu/sum(x) >= 0.5) {
      return(y[i])
      }
    else {
      i = i + 1
      }
  }
}

# Median for weighted temp difference:
temp_median = recs %>% 
  filter(TEMPNITE >= 0 , EQUIPMUSE >= 0 , TEMPHOME >= 0) %>%
  mutate(equipuse = decode_all_equipmuse(EQUIPMUSE)) %>%
  select(TEMPHOME, TEMPNITE, weight = NWEIGHT, equipuse) %>%
  group_by(equipuse) %>%
  summarize(mean_median = my_median(weight, (TEMPHOME - TEMPNITE)))

repl_median = recs %>% 
  filter(TEMPNITE >= 0 , EQUIPMUSE >= 0 , TEMPHOME >= 0) %>%
  gather(key = 'replicate', value = 'BRRWT', BRRWT1:BRRWT96) %>%
  mutate(equipuse = decode_all_equipmuse(EQUIPMUSE)) %>%
  select(TEMPHOME, TEMPNITE, BRRWT, replicate, equipuse) %>%
  group_by(equipuse,replicate) %>%
  summarize(rep_median = my_median(BRRWT, (TEMPHOME - TEMPNITE))) %>%
  ungroup() %>%
  left_join(temp_median, by = c('equipuse')) %>%
  group_by(equipuse) %>%
  summarize( mean_median = mean_median[1],
             std_err =  sqrt( mean( {rep_median - mean_median}^2 )*4/96 )) %>%
  mutate( lwr = mean_median - qnorm(.975)*std_err,
          upr = mean_median + qnorm(.975)*std_err
  ) %>%
  select(-std_err) %>%
  knitr::kable(digits = 1, caption = paste('
                                     Median difference between the daytime 
                                      (with someone home) 
                                      and nighttime temperatures for 
                                      each level of"main heating equipment 
                                      household behavior".'), 
               col.names = c('Heating Equipment Behavior', 
                             'Difference', 
                             'Lower Bound', 'Upper Bound'))%>%
  kable_styling(bootstrap_options = "striped",  full_width = F)
repl_median
```

## Question 2.a

```{r}
# Data:
kh2017 <- KH2017_raw
source('ps2_q2_funcs.R')
```


```{r, echo = FALSE}
trans = function(dat) {
  dat = as_tibble(dat)
  dat = select(dat,subject_nr,count_trial,xpos_get_response, 
               ypos_get_response, timestamps_get_response)
  x = as.list(as.character(dat$xpos_get_response))
  x = str_sub(x,start = 2, end = (str_length(x)-1))
  x = str_split(x,', ')
  x = lapply(x, FUN = as.numeric)
  #similar for y and t
  y = as.list(as.character(dat$ypos_get_response))
  y = str_sub(y,start = 2, end = (str_length(y)-1))
  y = str_split(y,', ')
  y = lapply(y, FUN = as.numeric)
  t = as.list(as.character(dat$timestamps_get_response))
  t = str_sub(t,start = 2, end = (str_length(t)-1))
  t = str_split(t,', ')
  t = lapply(t, FUN = as.numeric)
  dattran = transmute(dat,subject_nr,count_trial,x,y,t)
  return(dattran)
}

```

## Question 2.b - c

```{r, echo = FALSE}
recsq2c <- filter(kh2017,correct == 1)
recsq2ctrans = trans(recsq2c)
n = nrow(recsq2c)
result1 = matrix(data = NA, nrow=n, ncol = 1)
result2 = matrix(data = NA, nrow=n, ncol = 1)
result3 = matrix(data = NA, nrow=n, ncol = 1)
result4 = matrix(data = NA, nrow=n, ncol = 1)

for (i in 1:n) {
  mat = cbind(unlist(recsq2ctrans$x[i]),
              unlist(recsq2ctrans$y[i]),
              unlist(recsq2ctrans$t[i]))
  mat = normalize_traj(mat)
  result1[i] = comp_dist(mat[,1],mat[,2])
  result2[i] = max( abs(mat[, 2]) )
  result3[i] = mean( abs(mat[, 2]) )
  result4[i] = comp_auc(mat[,1],mat[,2])
}
result4q2c = transmute(recsq2ctrans,subject_nr,count_trial, 
                  Condition = recsq2c$Condition,
                  Exemplar = recsq2c$Exemplar,
                  tot_dist = result1,
                  max_abs_dev = result2,
                  avg_abs_dev = result3,
                  AUC = result4) 
output = result4q2c %>%
  knitr::kable(digits = 1, format = 'markdown',
             caption = paste('Measures of curvature computed from the KH2017_raw data set.'))
output
```

## Question 2.d

```{r, echo = FALSE}
s0 = summary(lmer(log(tot_dist) ~ Condition + (1 | subject_nr) + (1 |Exemplar), data = result4q2c))$coefficients

summaryintotal0 = c(estimate = s0[,1],lower = s0[,1] - 2 * s0[,2],
                    upper = s0[,1] + 2 * s0[,2])
summaryintotal0 = matrix(summaryintotal0, ncol = 3, nrow = 2,
                         dimnames = list(c('Intercept','Condition'),
                                       c('Estimate_tot_dist', 'Lower_tot_dist',
                                         'Upper_tot_dist')))


s1 = summary(lmer(log(max_abs_dev) ~ Condition + (1 | subject_nr) + (1 | Exemplar),data = result4q2c))$coefficients
summaryintotal1 = c(estimate = s1[,1],lower = s1[,1] - 2 * s1[,2],
                    upper = s1[,1] + 2 * s1[,2])
summaryintotal1 = matrix(summaryintotal1,ncol = 3, nrow = 2,
                      dimnames = list(c('Intercept','Condition'),
                                      c('Estimate_max_abs_dev',
                                        'Lower_max_abs_dev',
                                        'Upper_max_abs_dev')))

s2 = summary(lmer(log(avg_abs_dev) ~ Condition + (1 | subject_nr) + (1 | Exemplar), 
    data = result4q2c))$coefficients
summaryintotal2 = c(estimate = s2[,1],lower = s2[,1] - 2 * s2[,2],
                    upper = s2[,1] + 2 * s2[,2])
summaryintotal2 = matrix(summaryintotal2, ncol = 3, nrow = 2,
                      dimnames = list(c('Intercept','Condition'),
                                      c('Estimate_avg_abs_dev',
                                        'Lower_avg_abs_dev',
                                        'Upper_avg_abs_dev')))

s3 = summary(lmer(log(abs(AUC)) ~ Condition + (1 | subject_nr) + (1 | Exemplar),data = result4q2c))$coefficients
summaryintotal3 = c(estimate = s3[,1], lower = s3[,1] - 2 * s3[,2],
                    upper = s3[,1] + 2 * s3[,2])
summaryintotal3 = matrix(summaryintotal3, ncol = 3, nrow = 2,
                      dimnames = list(c('Intercept','Condition'),
                                      c('Estimate_AUC','Lower_AUC','Upper_AUC')))

sumary1 <- cbind(summaryintotal0, summaryintotal1) 
sumary2 <- cbind(summaryintotal2, summaryintotal3)
sumary3 <- rbind(sumary1, sumary2)
knitr::kable(sumary3, digits = 1, caption = '*Fitness of different models per curvature measure with "Condition".*')  %>%
  kable_styling(bootstrap_options = "striped", font_size = 6, full_width = F) 
sumary3
```

Which one does condition have the largest (relative) effect?

From the summary above, we can find that avg_abs_dev has the largest relative effect.

Since the coefficient of Condition in that model has largest abs value. The larger departure from 0 means has bigger effect.
That is the same for confidence intervial. The CI for the third model has the largest departure from 0.


## Question 1.b

```{r, echo=FALSE}
# Using the decode functions above get a table in one-line:
fuel_pro = recs %>% 
  filter(HEATHOME == 1) %>%
  gather(key = 'replicate', value = 'BRRWT', BRRWT1:BRRWT96) %>%
  mutate(division = decode_all_division(DIVISION), 
         uatyp = decode_all_uatyp(UATYP10), 
         fuel = decode_all_fuel(FUELHEAT)) %>%
  select(division, uatyp, fuel, heathome = HEATHOME, 
         weight = NWEIGHT, replicate, BRRWT) %>%
  replace_na(list(weight = 0)) %>%
  group_by(division, uatyp, fuel, replicate) %>%
  summarize(repweight = sum(BRRWT), fuelpro = sum(weight)) %>%
  group_by(division, uatyp, replicate) %>%
  mutate(reppro = 100 * repweight / sum(repweight), 
        pro = 100 * fuelpro / sum(fuelpro)) %>%
  group_by(division, uatyp, fuel)  %>%
  summarize(pro = pro[1],
             std_err =  sqrt(mean({reppro - pro}^2)*4/96)) %>%
  mutate(lwr = pro - qnorm(.975)*std_err,
        upr = pro + qnorm(.975)*std_err,
        pro = sprintf("(%f,%f) ;\n %f",lwr,upr,pro) ) %>%
  select(-std_err, -lwr, -upr) %>%
  tidyr::pivot_wider(names_from = fuel, 
                     values_from = pro) %>%
  knitr::kable(digits = 1, caption = paste('Proprtion of fuel types by division and urban type.','Each row shows the fuel type distribution for one or more division and urban types.')) %>%
  kable_styling(bootstrap_options = "striped", font_size = 4, full_width = F) %>%
    kableExtra::landscape() 
fuel_pro
```